---
title: "Bayesian estimation of slope of grade scale"
author: "Alexei Drummond"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_notebook: default
  pdf_document: default
params:
  data.set.name: "Germany"
  data.set.name.short: "germany"
  endDate: '2021-08-01'
  startDate: '2016-08-01'
  filter.by.tickprop.lrp: FALSE
  gear: "Sport"
  grade.type: "UIAA"
  in.path: "/Users/alexeidrummond/Git/climbing-grades/data/processed/"
  min.ascents: 30
  min.failures: 1
  min.grade: "5+"
  mean.grade.prior: "6"
  max.climbers: 100
  out.path: "/Users/alexeidrummond/Git/climbing-grades/results/"
  per.session: FALSE
---

# Analysis `r params$gear` `r params$data.set.name`

The ascent data was obtained from [thecrag.com](thecrag.com) using the provided API [https://www.thecrag.com/en/article/api](https://www.thecrag.com/en/article/api). 

For this analysis the data was filtered to include ascents from `r params$data.set.name` between `r params$startDate` and `r params$endDate`, of grade type `r params$grade.type` and in the `r params$gear` style. The minimum grade considered was `r params$min.grade` and only climbers with at least `r params$min.ascents` ascents during the time period, and at least `r params$min.failures` explicit failures were included. `r if(params$per.session){"The data was filtered to session logging, so that only the best ascent of each route on each day for each climber was kept."}`

A summary of the data processing is shown in Table \@ref{tab:data-processing}.

```{r setup, include=FALSE}
  knitr::opts_chunk$set(error = TRUE, warning = TRUE)
```

```{r initial-setup, echo=FALSE}
setwd("/Users/alexeidrummond/Git/climbing-grades/R")

source("produce-analysis-data.R")
```


```{r read-in-data, echo=FALSE}
#########################################
# Analysis settings
#########################################

res <- readRDS(paste0(params$in.path,"ascents-",params$data.set.name.short,".rds"))
routes <- readRDS(paste0(params$in.path,"routes-",params$data.set.name.short,".rds"))
```

```{r produce-analysis-data, echo=FALSE}
data <- produce.analysis.data(params, res, routes)

#store params for later processing steps
data$params <- params
```

```{r heatmap, echo=FALSE}
t <- table(data$df$ascent.type, data$df$account.id)

#normalise by column sums
#t <- t/colSums(t)[col(t)]

heatmap(t)
```

```{r regression-plot, echo=FALSE}
include_graphics(data$regressionpng)
```

The average slope from a simple regression of the log-odds of failure is
```{r}
mean(data$m, na.rm=TRUE)
```
The average grade range from a simple regression of the log-odds of failure is
```{r}
summary(data$grade.range)
```

```{r data-processing-table, results = 'asis', echo=FALSE}
##########################################################################################
# Write filter results to table
##########################################################################################
caption <- paste0("Summary of data processing for analysis of ", params$data.set.name, " ascent data.")

label <- paste0("table-data-processing-",params$data.set.name.short);

xt <- xtable(data$filter, caption = caption, label = label)

tablefile <- paste0(data$file.stem, "-processing-table.tex")

print(xt, 
      size="\\fontsize{9pt}{10pt}\\selectfont", 
      include.rownames=FALSE,
      sanitize.colnames.function = function(x) {paste0("{\\bf ",x,"}")}, 
      file=tablefile)

kable(data$filter, caption=caption, label=label)
```

The final processed data set includes a total of 

```{r data-summary, echo=FALSE, result='as is' }
data.summary <- data.frame(label=c("Climbers", "Ascents","Total Months"), value=c(data$d$C, data$d$N, data$d$P))

kable(data.summary)
```

Not all climbers made ascents during all months. 

The total number of climber months is
```{r}
sum(data$d$maxPage - data$d$minPage + 1)
```

The average number of ascents per climber per active month is
```{r}
data$d$N / sum(data$d$maxPage - data$d$minPage + 1)
```

```{r stan-analysis, echo=FALSE}
##########################################################################################
# RUN BAYESIAN INFERERENCE
##########################################################################################

stan.result <- run.stan.climbing.model(data$d)

fit1 <- stan.result$fit
data$time <- stan.result$time
```


```{r bayesian-analysis-figure, echo=FALSE}
 pngname <- paste0(data$file.stem, "-posterior.png")
 
 ylab <- "Flash Grade";
 if (params$per.session) {
   ylab <- "Session Grade"
 }
 
 plot.stan.climbing.results(params$startDate, params$endDate, fit1, data$d, pngname, ylab=ylab, data$df);
 
 include_graphics(pngname)
```


```{r save-results, echo=FALSE}
fit.file.name <- paste0(data$file.stem, ".rds")

saveRDS(as.data.frame(fit1), file=fit.file.name)
rm(fit1)

data.file.name <- paste0(data$file.stem, "-data.rds")
saveRDS(data, file=data.file.name)
```

```{r load-results, echo=FALSE}
fit.file.name <- paste0(data$file.stem, ".rds")

df <- readRDS(file=fit.file.name)

```

```{r mean-posterior, echo=FALSE}

q <- quantile(df$m, c(0.025, 0.5, 0.975))

print(q)

print(exp(q))
```
```{r mean-posterior-grades, echo=FALSE}

mean.grade <- compute.mean.grades(df, data$d$C, data)

hist(mean.grade)

hist(data$m)

```

```{r}
par(mar=c(3,7,1,1))

tab <- sort(table(data$df$ascent.type))

col <- ifelse(rownames(tab) %in% clean.ascent.types(), "green", "gray")

barplot(tab, horiz=T, col=col, las=1)
```


```{r}
summary  <- table(data$df$account.id, data$df$ascent.type)

explicit.fail.cols <- colnames(summary)[colnames(summary) %in% explicit.failed.ascent.types()]

summary(rowSums(summary[,explicit.fail.cols]))

fail.fraction <- rowSums(summary[,explicit.fail.cols])/rowSums(summary)

summary(fail.fraction)

```

```{r}

if (length(data$m) == length(fail.fraction) && sum(is.na(data$m))==0) {

  plot(data$m, fail.fraction, pch=19, col="blue")
  lm <- lm(fail.fraction~data$m)
  abline(lm, col="red")
}
```